### 2010-06-10 code/subarchitectures/visualization.sa/trunk/doc/v11n.tex
v11n: docs updated

\begin{itemize}
   \item Add visualization.sa to SVN externals:
   \begin{verbatim}
      $ svn pe svn:externals .
   \end{verbatim}
   Add the line:
   \begin{verbatim}
      ^/code/subarchitectures/visualization.sa/trunk   subarchitectures/visualization.sa
   \end{verbatim}

   % TODO: the script may be in another location
   \item install the required libraries; there is a script in visualization.sa/config
   that you can run:
   \begin{verbatim}
      sudo ubuntu-install-deps.sh
   \end{verbatim}

@@

The library provieds a display server that displays objects created by its
clients.  The objects that are on the display server are groupped into views.
Initially every new object is displayed in its own view that has the same name
as the object.
% (TODO: Additional views can be defined on the server that display multiple objects).
The display server creates rendering contexts in which the views are displayed.
There are two types of contexts: 2D and OpenGL.
% (TODO: new contexts: tabular; textual; ...)

The client that wants to display an object connects to the server and sends the
serialized object to the server. Each object has a name. If an object with the
same name already exists on the srver it is updated with the new data. Some
objects are composed of multiple parts where each part can be manipulated
independently.

Currently the following types of objects are suported: images, SVG graphics and
OpenGL objects define with LuaGl script or as TomGine render-models.

@@

The File system monitor is a CAST component that monitors the filesystem for
changes. Usualy it observes a directory on the disk and waits for files in that
directory that have been closed after they were modified. If the modified file
is one of the files the monitor is waiting for, it reads the file processes it
and sends it to the server. A single component is able to monitor multiple files
in multiple directories.

@@

\begin{verbatim}
   HOST localhost
   SUBARCHITECTURE visualization.sa 
   CPP WM SubarchitectureWorkingMemory
   CPP TM AlwaysPositiveTaskManager

@@

   CPP MG fileviewer.cli FileMonitorDC --display-server "display.srv"
      --setvars "dir=subarchitectures/visualization.sa/src/c++/clients/fileviewer/test/xdata"
      --monitor "
         CurrentPlan=dot:%(dir)/{*.dot}
         AltPlan=twopi:%(dir)/{*.dot}
         ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
         SvgFile=%(dir)/{*.svg}
         %(dir)/{*}"
         --log true --debug true
\end{verbatim}

The file monitor first connects to the server using the parameter
\code{--display-server}. If the server is not present, a display client will
operate normally except it won't be able to display its resutls.

The parameter \code{--setvars} sets some variables thata are used later when
the component configures the monitored files. The format of the parameter
string is \code{"var1=value1 var2=value2 ..."}. Later in the paramteter
\code{--monitor} a variable can be inserted with \code{\%(var1)}.

The most important and complex part is the definition of the files to be
monitored, the \code{--monitor} parameter. It is composed of multiple entries
that are separated by spaces. Each entry defines the directory where the files
are, the names of the files to monitor, the converters for the files and the
name of the object that will be created from the files. The general format of
each entry is:
\begin{verbatim}
   ObjectName=converter:directory/{filemasks}
\end{verbatim}

@@

\begin{verbatim}
   CurrentPlan=dot:%(dir)/{*.dot}
   AltPlan=twopi:%(dir)/{*.dot}
\end{verbatim}

@@

\begin{verbatim}
   ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
   SvgFile=%(dir)/{*.svg}
\end{verbatim}

@@

\begin{verbatim}
   SvgFile=svg:%(dir)/{*.xml}
\end{verbatim}

@@

\begin{verbatim}
   %(dir)/{*}"
\end{verbatim}

@@

\begin{verbatim}
   include_directories(${VISUALIZATION_INCLUDE_DIRS})
   link_cast_component(${CAST_COMPONENT_NAME} Video ${VISUALIZATION_LIBRARIES})
\end{verbatim}

@@

\begin{verbatim}
   add_definitions(-DFEAT_VISUALIZATION_OPENCV)
\end{verbatim}

Now the component is ready to interact with the display server. The component
will communicate with the servet through a display client:

\begin{verbatim}
   class VideoViewer : public ManagedComponent, public VideoClient
   {
   private:
   #ifdef FEAT_VISUALIZATION
     cogx::display::CDisplayClient m_display;
   #endif
   };
\end{verbatim}

Note that all the code related to visualization.sa is enclosed in conditional
blocks. The variable \code{FEAT\_VISUALIZATION} is defined when
viusalisation.sa is being built. If the visualization.sa is not used in the
project, the componetns that depend on it will continue to work, but they won't
display their data unless an alternative is provided (see VideoViewer.cpp which
uses OpenCV when visualization.sa is not used).

@@

\begin{verbatim}
   void VideoViewer::configure(const map<string,string> & _config)
   {
   #ifdef FEAT_VISUALIZATION
     m_display.configureDisplayClient(_config);
   #endif
   }
\end{verbatim}

The connection between the client and the server is established in the component's
\code{start()} method:
\begin{verbatim}
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
      m_display.connectIceClient(*this);
   #endif
   }
\end{verbatim}

@@

\begin{verbatim}
   void VideoViewer::receiveImages(const std::vector<Video::Image>& images)
   {
   #ifdef FEAT_VISUALIZATION
     m_display.setImage(getComponentID(), images[0]);
   #endif
   }
\end{verbatim}

@@

\begin{verbatim}
   class VideoViewer : public ManagedComponent, public VideoClient
   {
   private:
   #ifdef FEAT_VISUALIZATION
     cogx::display::CActiveDisplayClient<VideoViewer> m_display;
     void handleGuiEvent(const Visualization::TEvent &event);
     std::string getControlState(const std::string& ctrlId);
   #endif
   };
\end{verbatim}

@@

\begin{verbatim}
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
     m_display.connectIceClient(*this);
     m_display.installEventReceiver();
     m_display.setEventCallback(this, &VideoViewer::handleGuiEvent);
     m_display.setStateQueryCallback(this, &VideoViewer::getControlState);
   #endif
   }
\end{verbatim}

We define a checkbox element that will start or stop the flow of images
from the server:

\begin{verbatim}
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
   #endif
   }
\end{verbatim}

Now all that is missing are the callback functions:

\begin{verbatim}
   #ifdef FEAT_VISUALIZATION
   void VideoViewer::handleGuiEvent(const Visualization::TEvent &event)
   {
     if (event.type == Visualization::evCheckBoxChange) {
       if (event.sourceId == "toggle.viewer.running") {
         bool newrcv = (event.data != "0");
         if (newrcv != receiving) {
           if(receiving) {
             videoServer->stopReceiveImages(getComponentID());
           }
           else {
             videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
           }
           receiving = !receiving;
         }
       }
     }
   }
   #endif
\end{verbatim}

\begin{verbatim}
   #ifdef FEAT_VISUALIZATION
   std::string VideoViewer::getControlState(const std::string& ctrlId)
   {
     if (ctrlId == "toggle.viewer.running") {
       return receiving ? "2" : "0";
     }
     return "";
   }
   #endif
\end{verbatim}

>>>>

\begin{itemize}
   \item Add visualization.sa to SVN externals:
   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
      $ svn pe svn:externals .
   \end{Verbatim}
   Add the line:
   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
      ^/code/subarchitectures/visualization.sa/trunk subarchitectures/visualization.sa
   \end{Verbatim}

   % TODO: the script may be in another location
   \item install the required libraries; there is a script in visualization.sa/config
   that you can run:
   \begin{Verbatim}[fontsize=\scriptsize,gobble=6]
      sudo ubuntu-install-deps.sh
   \end{Verbatim}

@@

The library provides a display server that displays objects created by its
clients.  The objects that are on the display server are grouped into views.
Initially every new object is displayed in its own view that has the same name
as the object.
\todo{(TODO: Additional views that display multiple objects can be defined on the server)}
The display server creates rendering contexts in which the views are displayed.
There are two types of contexts: 2D and OpenGL.
% (TODO: new contexts: tabular; textual; ...)

The client that wants to display an object connects to the server and sends the
serialized object to the server. Each object has a name. If an object with the
same name already exists on the server it is updated with the new data. Some
objects are composed of multiple parts where each part can be manipulated
independently.

Currently the following types of objects are supported: images, SVG graphics and
OpenGL objects define with LuaGl script or as TomGine render-models.

@@

The File system monitor is a CAST component that monitors the file system for
changes. Usually it observes a directory on the disk and waits for files in that
directory that have been closed after they were modified. If the modified file
is one of the files the monitor is waiting for, it reads the file processes it
and sends it to the server. A single component is able to monitor multiple files
in multiple directories.

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   HOST localhost
   SUBARCHITECTURE visualization.sa 
   CPP WM SubarchitectureWorkingMemory
   CPP TM AlwaysPositiveTaskManager

@@

   CPP MG fileviewer.cli FileMonitorDC --displayserver "display.srv"
      --setvars "dir=subarchitectures/visualization.sa/src/c++/clients/fileviewer/test/xdata"
      --monitor "
         CurrentPlan=dot:%(dir)/{*.dot}
         AltPlan=twopi:%(dir)/{*.dot}
         ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
         SvgFile=%(dir)/{*.svg}
         %(dir)/{*}"
         --log true --debug true
\end{Verbatim}

The file monitor first connects to the server using the \code{--displayserver}
parameter. If the server is not present, a display client will operate normally
except it won't be able to display its results.

The parameter \code{--setvars} sets some variables that are used later when
the component configures the monitored files. The format of the parameter
string is \code{"var1=value1 var2=value2 ..."}. Later in the parameter
\code{--monitor} a variable can be inserted with \code{\%(var1)}.

The most important and complex part is the definition of the files to be
monitored, the \code{--monitor} parameter. It is composed of multiple entries
that are separated by spaces. Each entry defines the directory where the files
are, the names of the files to monitor, the converters for the files and the
name of the object that will be created from the files. The general format of
each entry is:
\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   ObjectName=converter:directory/{filemasks}
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   CurrentPlan=dot:%(dir)/{*.dot}
   AltPlan=twopi:%(dir)/{*.dot}
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   ImageFile=%(dir)/{*.png;*.jpg;*.pnm;*.xpm}
   SvgFile=%(dir)/{*.svg}
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   SvgFile=svg:%(dir)/{*.xml}
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   %(dir)/{*}"
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   include_directories(${VISUALIZATION_INCLUDE_DIRS})
   link_cast_component(${CAST_COMPONENT_NAME} Video ${VISUALIZATION_LIBRARIES})
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   add_definitions(-DFEAT_VISUALIZATION_OPENCV)
\end{Verbatim}

Now the component is ready to interact with the display server. The component
will communicate with the server through a display client:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   class VideoViewer : public ManagedComponent, public VideoClient
   {
   private:
   #ifdef FEAT_VISUALIZATION
     cogx::display::CDisplayClient m_display;
   #endif
   };
\end{Verbatim}

Note that all the code related to visualization.sa is enclosed in conditional
blocks. The variable \code{FEAT\_VISUALIZATION} is defined when
visualization.sa is being built. If the visualization.sa is not used in the
project, the components that depend on it will continue to work, but they won't
display their data unless an alternative is provided (see VideoViewer.cpp which
uses OpenCV when visualization.sa is not used).

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void VideoViewer::configure(const map<string,string> & _config)
   {
   #ifdef FEAT_VISUALIZATION
     m_display.configureDisplayClient(_config);
   #endif
   }
\end{Verbatim}

The connection between the client and the server is established in the component's
\code{start()} method:
\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
      m_display.connectIceClient(*this);
   #endif
   }
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void VideoViewer::receiveImages(const std::vector<Video::Image>& images)
   {
   #ifdef FEAT_VISUALIZATION
     m_display.setImage(getComponentID(), images[0]);
   #endif
   }
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   class VideoViewer : public ManagedComponent, public VideoClient
   {
   private:
   #ifdef FEAT_VISUALIZATION
     cogx::display::CActiveDisplayClient<VideoViewer> m_display;
     void handleGuiEvent(const Visualization::TEvent &event);
     std::string getControlState(const std::string& ctrlId);
   #endif
   };
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
     m_display.connectIceClient(*this);
     m_display.installEventReceiver();
     m_display.setEventCallback(this, &VideoViewer::handleGuiEvent);
     m_display.setStateQueryCallback(this, &VideoViewer::getControlState);
   #endif
   }
\end{Verbatim}

The flow of images from the VideoServer to the VideoViewer can be stopped at
any time. To let the user control the flow the component has to add a UI
element to the display server.  An appropriate type of UI element for this
purpose is a chekcbox:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void VideoViewer::start()
   {
   #ifdef FEAT_VISUALIZATION
     m_display.addCheckBox(getComponentID(), "toggle.viewer.running", "&Streaming");
   #endif
   }
\end{Verbatim}

The first parameter is the name of the {\em view} that will display the UI
element when it is active, not the name of object. This means that the UI
element will be displayed only when an object is displayed in its default view.
When an object is displayed in another view, the UI element won't be available.
\todo{TODO: user-defined views will be able to specify which elements to show}

The \code{handleGuiEvent()} callback function will be called every time the
UI element emits a known event. For the chekcbox element we need to respond
to the \code{evCheckBoxChange} event:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   #ifdef FEAT_VISUALIZATION
   void VideoViewer::handleGuiEvent(const Visualization::TEvent &event)
   {
     if (event.type == Visualization::evCheckBoxChange) {
       if (event.sourceId == "toggle.viewer.running") {
         bool newrcv = (event.data != "0");
         if (newrcv != receiving) {
           if(receiving) {
             videoServer->stopReceiveImages(getComponentID());
           }
           else {
             videoServer->startReceiveImages(getComponentID(), camIds, 0, 0);
           }
           receiving = !receiving;
         }
       }
     }
   }
   #endif
\end{Verbatim}

Note that the data for UI elements passed between the client and the server is
serialized into strings. The checkbox element knows only 2 values: \code{"0"}
and everything else.
\todo{TODO: Add more element types, define values}

A UI element may be displayed in multiple views. When a UI element is first
displayed it must be put in a state that reflects the current state of the
component.  Since a UI element may be created at any time it wouldn't be enough
to set the initial state of the element at creation time. Instead, we define
the \code{getControlState()} callback function:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   #ifdef FEAT_VISUALIZATION
   std::string VideoViewer::getControlState(const std::string& ctrlId)
   {
     if (ctrlId == "toggle.viewer.running") {
       return receiving ? "2" : "0";
     }
     return "";
   }
   #endif
\end{Verbatim}

When the function returns \code{""} the display server won't change the state
of the UI element.

\section{Displaying objects}

The server can display various types of objects. Initially it was created to
display images, but with time other objects (supported by Qt) were added:
SVG images and OpenGL scenes.

\subsection{Images}

The display client can send various types of images to the server: raw images, 
\code{Video::Image}, OpenCV's \code{IplImage} and encoded images that are
supported by Qt. An \code{IplImage} can only be sent if OpenCV was enabled
during compilation.
The clients interface for sending images is:

\begin{Verbatim}[fontsize=\scriptsize,gobble=3]
   void setImage(const std::string& id, int width, int height, int channels,
     const std::vector<unsigned char>& data);
   void setImage(const std::string& id, const Video::Image& image); 
   void setImage(const std::string& id, const std::vector<unsigned char>& data,
      const std::string &format=""); 
   #ifdef FEAT_VISUALIZATION_OPENCV
      void setImage(const std::string& id, const IplImage* pImage); 
   #endif
\end{Verbatim}

Raw images can have 1 or 3 channels and the specified image size must match the
size of \code{data}. The supported formats of the compressed image are: bmp,
gif, jpeg, jpg, png, pbm, pgm, ppm, tiff, xbm, xpm. It the format is not
specified, Qt will try to deduce the type from \code{data}. An \code{IplImage}
is first converted to a contiguous array of bytes and sent to the server as a
raw image.

Every image has an id which uniquely identifies an image on the server. This way
an image (as any other object) can be replaced on the server.
